Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis

### 2021-07-19 analysis/chapters/paper_ana_20181122.Rmd
Algunas correcciones menores.

---
title: "Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated. We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November, 22th, 2018. The accumulated precipitation observed during the transit of the MCS was quite high with values over 200 mm over the northern part of the domain. This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov. 2018.

@@

Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels. 
Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
@miyoshi2012a assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation. 
Moreover @bao2015 studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. 
Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].

@@

We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis. Since the publicly available GFS ensemble members change at each cycle **(alguna fuente)** we used random perturbations to generate the initial ensemble perturbations. This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015]. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in [@necker2020; @maldonado2021]. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configurations consisting on the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). All ensemble members used the same land-surface model (Noah-MP, @chen2001) and microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

@@

The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) **[ver cita]**. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts. The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment.  The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.

Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of this type of observation (99 % of observation).

We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are part of RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.

Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors. The observation errors were specified following the GSI default configuration. In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind). In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)). Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. This is evident in Figure \@ref(fig:obs-cycle)a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiments. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have a good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.

The channels accepted for assimilation and their associated error were defined using the default GSI configuration. The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, no the observation quality (based on available data quality predictors) and on the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred from those over land or snow [@hu2018].

@@

$$\mathit{BT_bc} =\mathit{ BT} + \sum{i = 0}{N} \beta_i p_i (x)$$

GSI has a constant offset bias correction term ($p_0 = 1$). The following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity. Scan angle-dependent bias is modelled as a 4th-order polynomial [@zhu2014].

The $\beta_i$ coefficients are usually trained to achieve the best fit between the simulation and the observations. In the GSI system these coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations. The coefficients initialized from the 18 UTC Nov, 11 2018 GFS coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments  in the estimated coefficients at each cycle.

In our experiments, only clear-sky observations are used. For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes[@weston2019; @zhu2016]. For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms. Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60$^{\circ}$) to reject channels near the visible range that can be contaminated with reflection. It also performs an emissivity check for observations over land for both infrared and microwave radiances.

@@

(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle and divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.

@@

* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements. Only the measurements/soundings that did not enter the assimilation system were used for validation. The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b.

* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill 1-hour forecasts to represent the precipitation over the domain.

@@

The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal). The larger number of assimilated observations over the center and east of the domain correspond to the ASWS observations. In this region we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet. In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown. Local maxima at 12 and 00 UTC  found mainly in CONV is produced by the operational soundings. The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage. The maxima after 12 and 00 UTC correspond to the contribution of the multispectral sensors. The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b shows a maximum in low levels due to the ASWS observations. Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa). Above 850 hPa, most of the observations correspond to radiance observations.

@@

```{r table-exp}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "PREPBUFR", "x", "x", "x", "x",
  "AUTSFC", " ", "x", "x", "x",
  "SATWND", " ", " ", "x", "x",
  "RAD", " ", " ", " ", "x",
) %>% 
  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      align = "lcccc",
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = TRUE) %>% 
  column_spec(1, width = "5em") %>% 
  column_spec(2:5, width = "3.5em")
```

@@

A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper. These aspects include a validation of how uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated . 
To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:

@@

where $x_o$ is the error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

where $M$ is the sample size. A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1, $sd RCRV = 1$. If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative). A $sd RCRV > 1$ ($sd RCRV < 1$), indicates that the ensemble is underdispersive (overdispersive).

@@

where ${P_x}$ and is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise. This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values. The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
  #                   proj = norargentina_lambert, color = "darkorange", 
  #              breaks = c(-10, -5), skip = 0,
  #              label.placement = label_placement_n(1)) +
  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1}")) +
  #cordillera +
  geom_mapa() +
  facet_grid(. ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        # legend.box.spacing = margin(0, 0, 0, 0),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +

>>>>

---
title: "Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  The impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF is evaluated. We conducted a case study corresponding to a huge mesoscale convective system (MCS) developed over central and north-eastern Argentina during November, 22th, 2018. The accumulated precipitation observed during the transit of the MCS was quite high, exceeding 200 mm over the northern part of the domain. This MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign during Nov. 2018.

@@

Given the sparsity of the local conventional observing network it is important to explore the potential benefits of including additional sources of observations at a regional scale. 
Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e. AIRS and IASI) in a regional domain application. 
@miyoshi2012a assimilated observations, atmospheric temperature and humidity profile data from the AIRS and found improvement on the tracking of two typhoons. 
@dillon2019 also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation. 
Moreover @bao2015 studied the impact on temperature and humidity forecast over western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. Recently @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed a positive improvement for all variables, in particular for relative humidity at upper levels.
Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts [@gao2015].

@@

We use a 60-member ensemble where the ensemble mean at the beginning of the data assimilation cycle is initialized using the GFS deterministic analysis. Since the publicly available GFS ensemble members change at each cycle **(alguna fuente)** we used random perturbations to generate the initial ensemble perturbations. This method also helps to prevent an underestimation of the ensemble spread [@ouaraini2015]. The perturbations are generated as scaled differences between two random atmospheric states obtained from the Climate Forecast System Reanalysis (CFSR) data with 0.5$^{\circ}$ horizontal grid spacing with a smooth time evolution as in @necker2020 and @maldonado2021. In this way, we preserved the nearly hydrostatic and geostrophic equilibrium of larger scales. These random perturbations are also applied at the boundaries in order to keep the ensemble spread within the domain.

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configurations consisting on the combination of 3 moist convection schemes (Kain–Fritsch [@kain2004], Grell–Freitas [@grell2013], and Betts–Miller–Janjic [@janjic1994]) and 3 planetary boundary layer schemes (Yonsei University Scheme [@hong2006], Mellor–Yamada–Janjic Scheme [@janjic1994], and Mellor–Yamada Nakanishi Niino [@nakanishi2009]). All ensemble members used the same land-surface model (Noah-MP, @chen2001), microphysics (WRF single-moment 6–class squeme [@hong2006a]), and radiation processes (RRTMG shortwave and longwave scheme [@iacono2008]) parameterizations.

@@

The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS) **[ver cita]**. We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircrafts. The orange triangles in Figure \@ref(fig:dominio)a, are the surface stations and ship locations included in this experiment. The orange triangles in Figure \@ref(fig:dominio) a), are the surface station locations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.

Satellite-derived wind observations also come from GDAS and includes information from GOES-16 (estimated from the visible, infrared and water vapour channels) and METEOSAT 8 and 11 (visible and water vapour channels). Due to the domain covered by each of these satellites, GOES-16 is the primary source of satellite-derived winds (99 % of the observations).

We used additional conventional observations from 866 Automatic Surface Weather Stations (ASWS) that are part of  various public and private automatic station networks over Southern South America and are included in RELAMPAGO Data Set repository [@garcia2019]. These stations are indicated as green squares in Figure \@ref(fig:dominio)a. They have higher spatial coverage than the CSWS and a sampling frequency of 10 minutes in most cases. The variables measured by the stations varied between networks; all stations measure temperature, but only 395 stations provide humidity, 422 provide pressure and 605 provide wind information.

Table \@ref(tab:table-obs) lists all the observation types (i.e., pressure, temperature, specific humidity and wind) available for each source, together with their associated errors. The observation errors were specified following the GSI default configuration. In some cases, the error varies with height and it depends on the specific platform (aircraft and satellite-derived wind). In terms of quality control, a gross check was performed by the observation operator by comparing the innovation (the difference between the observation and the model-simulated observation based on the first-guess) with a predefined threshold (also included in Table \@ref(tab:table-obs)).

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)). Since the regional domain is located in the mid-latitudes and all satellite platforms are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. This is evident in Figure \@ref(fig:obs-cycle)a which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 % of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have a good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top model (50 hPa) used in the experiments.

The channels accepted for assimilation and their associated error were defined using the default GSI configuration. The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. Firstly a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, on the observation quality (based on available data quality information) and on the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred from those over land or snow [@hu2018].

@@

$$\mathit{BT_{bc}} =\mathit{ BT} + \sum_{i = 0}^{N} \beta_i p_i (x)$$

GSI has a constant offset bias correction term ($p_0 = 1$) and the following predictors are the cloud liquid water content (CLW), the temperature lapse rate at the pressure of maximum weight, the square of the temperature lapse rate at the pressure of maximum weight, and the emissivity sensitivity. Scan angle-dependent bias is modelled as a 4th-order polynomial [@zhu2014].

In the GSI system the $\beta_i$ coefficients are trained using a variational estimation method which solves for the $\beta_i$ that provides the best fit between the simulation and the observations. The coefficients initialized from the 18 UTC Nov, 11 2018 GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.

In our experiments, only clear-sky observations are used. For microwave radiances, observations potentially contaminated by clouds are detected using the scattering and Liquid Water Path (LWP) indexes [@weston2019; @zhu2016]. For the infrared channels cloud contaminated observations are detected using the transmittance profile calculated within the CRTM algorithms. Additionally, the GSI quality control for infrared sensors looks for observations over water with a large zenithal angle (over 60$^{\circ}$) to reject channels near the visible range that can be contaminated with reflection. It also performs an emissivity check for observations over land for both infrared and microwave radiances.

@@

(ref:obs-cycle) a) Number of available observations per cycle and b) time averaged number of observations per cycle divided into 50 hPa-depth vertical layers for the CONV (blue squares and line), AUT (light blue dots and line), SATWND (orange triangles and line) and RAD (red diamond and line) experiments.

@@

* The Multi-Network Composite Highest Resolution Radiosonde Data [@sondeos] from the RELAMPAGO field campaign database consisting of high resolution radiosondes launched from several locations during the Intensive Observation Periods (IOP) along with the operational measurements. Only the soundings that did not enter the assimilation system were used for validation. The experiment period covers IOP missions 7 and 8 during which 74 radiosondes were launched in a small area near the center of the experimental domain (Figure \@ref(fig:dominio)b).

* The Satellite precipitation estimation IMERG Final Run with 0.01$^{\circ}$ spatial resolution and 30 minutes temporal resolution [@huffman2018] was used as a reference state to validate the skill of 1-hour forecasts to represent the precipitation over the domain.

@@

The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \@ref(fig:obs-horizontal). The larger number of assimilated observations over the center and east of the domain correspond to the ASWS observations. In this region we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet. In Figure \@ref(fig:obs-cycle)a the number of assimilated observations over time is shown. Local maxima at 12 and 00 UTC  found mainly in CONV is produced by the operational soundings. The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage. The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors. The vertical distribution of the mean number of observations per cycle (Figure \@ref(fig:obs-cycle)b) shows a maximum in low levels due to the ASWS observations. Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa). Above 850 hPa, most of the observations correspond to radiance observations.

@@

```{r table-exp}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "PREPBUFR", "x", "x", "x", "x",
  "ASWS", " ", "x", "x", "x",
  "SATWND", " ", " ", "x", "x",
  "RAD", " ", " ", " ", "x",
) %>% 
  kbl(booktabs = TRUE, col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      align = "lcccc",
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = TRUE) %>% 
  column_spec(1, width = "5em") %>% 
  column_spec(2:5, width = "3.5em")
```

@@

A set of metrics are selected to evaluate different aspects of the analysis obtained in the experiments conducted in this paper. These aspects include a validation of how the uncertainty is quantified in the first-guess and in the analysis and how different experiments fit an independent set of observations which are not assimilated.

To evaluate the consistency of the uncertainty quantification in the first-guess and in the analysis we use the Reduced Centered Random Variable (RCRV, @candille2007) which is defined as:

@@

where $x_o$ is the observation and it error $\sigma_o$, the mean $m$, and the standard deviation $\sigma$ of the ensemble. 
The average of $RCRV$ computed over all the realizations represents the bias of the ensemble mean with respect to the observations normalized by the estimated uncertainty:

@@

where $M$ is the sample size. A reliable system will have no bias ($mean RCRV = 0$), and a standard deviation equal to 1 ($sd RCRV = 1$). If the ensemble has a positive (negative) bias, $mean RCRV$ will be positive (negative). A $sd RCRV > 1$ ($sd RCRV < 1$), indicates that the ensemble is underdispersive (overdispersive).

@@

where ${P_x}$ is the probability of occurrence of accumulated rainfall over a specified threshold computed from the ensemble and ${P_o}$ is 1 when the observed accumulated precipitation is over the same threshold and 0 otherwise. This probabilistic formulation uses the information from the full ensemble thus avoiding the smoothing effect of the ensemble mean which can particularly affect localized extreme precipitation values. The FSS was applied to the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

@@

ana[bottom_top %between% c(1, 7), .(v_mean = mean(v)), 
    by = .(south_north, west_east, date, exp)] %>% 
  .[pw, on = .NATURAL] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = value, fill = stat(level_d)),
                    proj = norargentina_lambert) +
  colorspace::scale_fill_continuous_divergingx(super = ScaleDiscretised,
                                               palette = "PRGn", mid = 25,
                                               guide = guide_colorsteps(barwidth = 25,
                                                                        barheight = 0.5, 
                                                                        title.position = "left", 
                                                                        title.vjust = 1,
                                                                        frame.colour = "black")) +
  geom_contour2(aes(z = v_mean, size = stat(level)),
                proj = norargentina_lambert, color = "black", 
                breaks = c(-10, -5)) +
  scale_size_continuous(breaks = c(-10, -5), range = c(0.8, 0.4),
                        labels = c("-10", "-5")) +
  # geom_text_contour(aes(z = v_mean), size = 3, stroke = 0.05,
  #                   proj = norargentina_lambert, color = "darkorange", 
  #              breaks = c(-10, -5), skip = 0,
  #              label.placement = label_placement_n(1)) +
  # scale_linetype_manual(guide = NULL, values = c(2, 1)) +
  labs(fill = latex2exp::TeX("Precipitable\nwater ($Kgm^{-2}$)"), size = latex2exp::TeX("V ($ms^{-1})")) +
  #cordillera +
  geom_mapa() +
  facet_grid(. ~ exp, 
             labeller = labeller(exp = c(E2 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E8 = "RAD"))) +
  theme_minimal() +
  theme(legend.position = "bottom",
        # legend.box.spacing = margin(0, 0, 0, 0),
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3)) +
